---
title: "PML peer assignment writeup"
author: "Amrendra"
date: "14/09/2020"
output: html_document
---

## Introduction
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross-validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

## Data
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.


### Loading libraries
```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```

### Downloading the Data using the links
```{r, cache = T}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile, method="curl")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile, method="curl")
}
```  
### Extracting the initial details by reading the Data
After downloading the data from the data source, we can read the two csv files into two data frames.  
```{r, cache = T}
trainRaw <- read.csv("./data/pml-training.csv")
testRaw <- read.csv("./data/pml-testing.csv")
dim(trainRaw)
dim(testRaw)
```
Training data have 19622 rows(observations) and 160 columns(variables) and 
test data have 20 rows(observations) and 160 columns(variables)

### Cleaning of the data
Observations have lot of missing values and nonsnse variables. TO get rid of such variables and oservations we need to clean the data
```{r, cache = T}
sum(complete.cases(trainRaw))
```
Removing the columns that have NA as their missing values.
```{r, cache = T}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
```  
Next, Cleaning the variables that have low or no effect on the models.
```{r, cache = T}
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```
Now, the cleaned training data set contains 19622 observations and 53 variables, while the testing data set contains 20 observations and 53 variables. The "classe" variable is still in the cleaned training set.

### Data Splitting
Then, we can split the cleaned training set into a pure training data set (60%) and a validation data set (40%). We will use the validation data set to conduct cross validation in future steps if needed.  
```{r, cache = T}
set.seed(12345) # For reproducibile purpose
inTrain <- createDataPartition(trainCleaned$classe, p=0.60, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

## Data Modeling
We build a predictive model for activity recognition using **Random Forest** algorithm and also We will use **5-fold cross validation** when applying the algorithm.  
```{r, cache = T}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```

## Predicting for Test Data Set
Now, we apply the model to the original testing data set downloaded from the data source.  
```{r, cache = T}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
```  